---
title: "R Notebook"
output: html_notebook
---

---
title: "R Notebook"
authors: Taylor, Menna, Shaun, Joel
output: html_notebook
---

Midway data analysis memo link: https://docs.google.com/document/d/1j_19XNibAk0Modj5MFnaPfIz1M4JZ1dzW1s5KP2PIQM/edit?usp=sharing 

Load Libraries

```{r}
library(tidyverse)
library(corrr)
library(tools)
library(janitor)
library(lubridate)
library(ggplot2) #make sure you have ggplot 2 installed, not ggplot
options(scipen=999)
#install.packages("corrr")

```

Let's clean the date columns now.
```{r}
###asked chatgpt how to turn a character column into a date time column, gave it some examples from our code and int told me to do this:
wh_visitor_data <- read_csv("data/cleaned_OPENREFINE.csv") |> 
  clean_names()
wh_visitor_data <- wh_visitor_data |> 
  mutate(appt_end_date = mdy_hm(appt_end_date))

#We need to take the time out of appt end date since some say 11:59 and some say 23:59.
wh_visitor_data <- wh_visitor_data |>
  mutate(appt_end_date = as.Date(appt_end_date))

```

Let's take a quick look at the data now that we've changed these columns into dates.

```{r}
glimpse(wh_visitor_data)
```

Here are our new questions we'd like to answer: 
  1. What are the most popular times / dates / months for White House visitors and how do they correlate with important political events? 
  2. Has the amount of visitors fluctuated during Biden presidency (see if this correlates with rise in drop of approval rating)
  3. West Wing Tours are selective, how many are being taken? How many were taken on July 2 when they found traces of cocaine?
  4. Who has sponsored the most White House visitors - e.g. who is getting the most visitors? 
  5. What percent of visitors have just been labeled "office visitors"? If it's a significant percentage, that would seem to defeat the point of releasing these logs. Not much better than the Trump administration.
  6. How often are AI lobbyists and CEOs coming to visit during talks about regulating AI?

Immediate limitations: 1. There are a lot of NAs in here where dates should go. I think the most reliable date is the appointment end date, which means the times are probably useless for us. 2. Dates and times aren't always written the same way. Is that going to cause us problems? We should probably check with Derek about that. 3. We wanted to look at how many folks are labeled office visitor - we're gonna have to find all instances of that, they're not marked the same. 4. We had a question about west wing visits. I don't see any identifying data here to show which ones are west wing visits, although there are a lot of confusing column names and abbreviations. 5. There are a ot of confusing column names and abbreviations - we will need to dig deeper into that to find out what we can actually look at here. 6. I still really want to know what types of people are visiting. Could we at least match congress people up using the propublica dataset and see how many folks are not house or senate members? We may be able to find a dataset of foreign powers we can match up as well.

Let's see when the most popular day for visiting was. First we need to figure out which date column to use.

```{r}

wh_visitor_data |>
  group_by(toa) |> 
  summarize(count=n()) |>
  arrange(desc(count))
  
# note - since there are so many NA dates here, I think we should group by appt date.

wh_visitors_by_date <- wh_visitor_data |>
  group_by(appt_start_date) |>
  summarize(total_visits = n()) |>
  arrange(desc(total_visits))

### just kidding, there are a ton of NA dates still. maybe appt end date?

wh_visitors_by_end_date <- wh_visitor_data |>
  group_by(appt_end_date) |>
  summarize(total_visits = n()) |>
  arrange(desc(total_visits))

### this seems better

wh_visitors_by_end_date

```


```{r}
###derek asked if the NA dates were specific to certain offices. how do we figure that out?

wh_visitor_data |> filter(if_any(-toa, is.na)) |>
  group_by(meeting_loc) |>
  summarise(count = n())

```

Now that we know what columns we need, let's finish cleaning our data.

```{r}
wh_visitor_data <- wh_visitor_data |>
  select(-uin, -bdgnbr, -access_type, -appt_cancel_date, -last_updatedby, -post, -lastentrydate, -terminal_suffix, -caller_name_last, -caller_name_first, -caller_room, -column, -column2, -tod, -pod)
```

Question 1.  What are the most popular months and dates? What political events correlate with those?

```{r}

###mutate(toa_time = ymd_hms(toa) |> time()) we need to figure out how to separate out time from these two columns as well

wh_visitor_data |>
  group_by(toa) |> 
  summarize(count=n()) |>
  arrange(desc(count))
  
# note - since there are so many NA dates here, I think we should group by appt date.

wh_visitors_by_date <- wh_visitor_data |>
  group_by(appt_start_date) |>
  summarize(total_visits = n()) |>
  arrange(desc(total_visits))

### just kidding, there are a ton of NA dates still. maybe appt end date?

wh_visitors_by_end_date <- wh_visitor_data |>
  group_by(appt_end_date) |>
  summarize(total_visits = n()) |>
  arrange(desc(total_visits))

#That looks better. Let's use appt end date because it seems to be the date col with the most data and fewest NAs.

wh_visitors_by_end_date
```

Top 5 Dates: 

1. 06-22-2023: President Biden and Prime Minister Modi of the Republic of India in Joint Press Conference
2. 12-23-2022: President Biden passes a successful bipartisan funding bill
3. 04-26-2023:President Biden and President Yoon Suk Yeol of the Republic of Korea to commemorate the 70th anniversary of their alliance
4. 07-04-2023: Independence Day event at the White House hosted by President Biden and First Lady Jill Biden (recurring date with a big total of visitors)
5. 12-13-2022: President Biden signs the Respect for Marriage Act, legalizing gay marriage

Another important date: July 2, the day cocaine was found in the White House.

This is actually pretty interesting! It's somewhat surprising that the top date doesn't solely have to do with the United States itself, but I remember driving through Capitol Hill several times that week and seeing the huge impact (both good and bad) that Modi's visit had on the city. The majority of these highest visited dates correlate with important international visits, United States-specific holidays, and the passage of significant things into law.

It may be helpful to check out the names of folks that are visiting on these days for more context

```{r}

# 06-22-2023: President Biden and Prime Minister Modi of the Republic of India in Joint Press Conference

wh_visitor_data |>
  filter(str_detect(appt_end_date, "2023-06-22")) |>
  select(total_people, namelast, namefirst, visitee_namelast, visitee_namefirst) |>
  arrange(desc(total_people))

# 12-23-2022: President Biden passes a successful bipartisan funding bill

wh_visitor_data |>
  filter(str_detect(appt_end_date, "2022-12-23")) |>
  select(total_people, namelast, namefirst, visitee_namelast, visitee_namefirst) |>
  arrange(desc(total_people))

# 04-26-2023: President Biden and President Yoon Suk Yeol of the Republic of Korea to commemorate the 70th anniversary of their alliance

wh_visitor_data |>
  filter(str_detect(appt_end_date, "2023-04-26")) |>
  select(total_people, namelast, namefirst, visitee_namelast, visitee_namefirst) |>
  arrange(desc(total_people))

# 07-04-2023: Independence Day event at the White House hosted by President Biden and First Lady Jill Biden (recurring date with a big total of visitors)

wh_visitor_data |>
  filter(str_detect(appt_end_date, "2023-07-04")) |>
  select(total_people, namelast, namefirst, visitee_namelast, visitee_namefirst) |>
  arrange(desc(total_people))

# 12-13-2022: President Biden signs the Respect for Marriage Act, legalizing gay marriage

wh_visitor_data |>
  filter(str_detect(appt_end_date, "2022-12-13")) |>
  select(total_people, namelast, namefirst, visitee_namelast, visitee_namefirst) |>
  arrange(desc(total_people))

```

Notable information gathered from examining top 5 dates for visitors: 
1. 06-22-2023: The biggest groups (832) are visiting Air Force Chief Master Sgt. Edward Teleky, Air Force Band's drum major. It takes about 83 pages before you can find folks visiting the office or the president. A good amount of these names are South Asian sounding which is consistent with our research.
2. 12-23-2022: All of these groups were office visitors and were around 250 people 
3. 04-26-2023: The biggest groups (920) are visiting the president. It takes about 55 pages for another name to show up. That other name is Ed Teleky. Many of these last names are East Asian sounding, which is consistent with our research.
4. 07-04-2023: The biggset groups (1182) are visiting Assistant Director of White House Operations Bonnie Casillas. About 82 pages in we see that the rest of the folks are visiting the president.
5. 12-13-2022: All of these groups (2155) were visiting the president

We think it may actually be more interesting to look at data where groups were less than 100 because those may be attributed to big events or holidays, which we could have guessed. What are the most popular dates and months for regular folks?

```{r}
wh_visitor_data |>
  group_by(appt_end_date) |> 
  filter(total_people <100) |>
  summarize(count=n()) |>
  arrange(desc(count))
```

Top 5 Dates with groups smaller than 100: 

1. 12-23-2022: President Biden passes a successful bipartisan funding bill
2. 03-18-2023: Biden hosts Irish Taoiseach at White House for St. Patrick’s Day
3. 05-11-2023: This was a big day at The House because the Biden-Harris Administration made a lot of big decisions about Global Health, Environmental Protection and National Conservation, Immigration, and International/National Emergency Funding
4. 03-25-2023: President Biden and President Alberto Fernandez of Argentina meet at the White House
5. 05-12-2023: Continuation of the decisions made on 5/11, as well as the start of Infrastructure Week.

This is so interesting because, aside from 12/23, none of the dates repeat in the data without big organized groups. This shows that big organized groups frequently visit on holidays and days with big international repercussions, while more national-focused events and big decisions are primarily attended by individuals or folks in smaller groups. It's a bit surprsing that St. Patrick's Day is so high up there, but Biden is Irish Catholic so it makes sense.

It may be helpful to check out the names of folks that are visiting on these days as well for more context

```{r}

# Skipping 12-23-2022 because it's already about

# 03-18-2023: President Biden passes a successful bipartisan funding bill

wh_visitor_data |>
  filter(str_detect(appt_end_date, "2023-03-18")) |>
  select(total_people, namelast, namefirst, visitee_namelast, visitee_namefirst) |>
  arrange(desc(total_people))

# 05-11-2023: Biden-Harris Administration make big decisions regarding Global Health, Environmental Protection and National Conservation, Immigration, and International/National Emergency Funding

wh_visitor_data |>
  filter(str_detect(appt_end_date, "2023-05-11")) |>
  select(total_people, namelast, namefirst, visitee_namelast, visitee_namefirst) |>
  arrange(desc(total_people))

# 03-25-2023: President Biden and President Alberto Fernandez of Argentina meet

wh_visitor_data |>
  filter(str_detect(appt_end_date, "2023-03-25")) |>
  select(total_people, namelast, namefirst, visitee_namelast, visitee_namefirst) |>
  arrange(desc(total_people))

# 05-12-2023: Continuation of the decisions made on 5/11, as well as the start of Infrastructure Week

wh_visitor_data |>
  filter(str_detect(appt_end_date, "2022-05-12")) |>
  select(total_people, namelast, namefirst, visitee_namelast, visitee_namefirst) |>
  arrange(desc(total_people))

```

Notable information gathered from examining top 5 dates for visitors with groups of less than 100: 
2. 03-18-2023: Visitors here vary in terms of who they are visiting, including the vice president, the office, Deputy Social Secretary to the VP Amanda Trocola, Senior Advisor Andres Anzola, UN Assistant Secretary Pierce Davis, Program Examiner for the Office of Management and Budget Nora Stein and more. 
3. 05-11-2023: All groups here (206) are visiting the president
4. 03-25-2023: All groups here are office visitors -- most are about 255 but some are as little as 30
5. 05-12-2023:Visitors here vary greatly in terms of who they're visiting. The biggest groups (150) are visiting Peyton Schwartz, the special assistant to the social secretary and deputy director of arrangements. Others are visitng folks like Ed Teleky, the president, Deputy Assistants to the President Erika Moritsugu and Danielle Carnival, and more.

This data isn't the most useful but it's somewhat interesting!

2. Has the amount of visitors fluctuated during Biden presidency (see if this correlates with rise in drop of approval rating)

ANSWER: Visits in 2021 were super minimal. In 2022, visits were fairly sparse in the first half of the year, but ramped up in July and the following months. In 2023 there were far more visits to the White House - could be from a full post-pandemic "return to normal" or could be due to things gearing up ahead of election season? Similar to 2022, there was a jump in visits in summer months. However, this year June was the top month, because of PM Modi's visit. There's no correlation between visits and approval ratings.

```{r}
### Note - asked ChatGPT how to make the month column have labels and gave it my mutate code creating a new col with lubridate
#reading in approval ratings data from gallup
approval_ratings <- read_csv("data/biden_approval_gallup.csv", col_names = FALSE)

colnames(approval_ratings) <- c("poll_date", "rep_approval_pct", "ind_approval_pct", "dem_approval_pct")
  
approval_ratings <- approval_ratings |>
  separate(poll_date, into = c("year", "month", sep = "-")) |>
  select(year, month, rep_approval_pct, ind_approval_pct, dem_approval_pct)

approval_ratings <- approval_ratings |>
  mutate(date = paste(year, month, sep = " "))

#making a df with top months for visitors
wh_visitor_by_month <- wh_visitors_by_end_date |> 
 mutate(date_time = ymd(appt_end_date),
  month = month.abb[month(appt_end_date)])

wh_visitor_by_month <- wh_visitor_by_month |>
  group_by(month) |>
  summarise(total_visits = sum(total_visits))

#last one not really helpful bc didn't sort by year. Now doing the same with year and month
wh_visitors_month_year <- wh_visitors_by_end_date |> 
 mutate(date_time = ymd(appt_end_date),
  month = month.abb[month(appt_end_date)],
  year = year(date_time))

top_months_by_year <- wh_visitors_month_year |>
  group_by(year, month) |>
   summarise(total_visits = sum(total_visits))
```



```{r}
#not sure if that was useful. I think I need to join these dataframes. Asked chatgpt for help with this because I needed to join on the date column and didn't have one with year and month in it for top_months_by_year. Asked it: I want to create a date column that looks like this: 2022 Jan, 2022 Feb, 2022 March. Currently I have a year column that looks like this: 2022, 2022, 2022, and a month column that looks like this: Jan, Feb, March. How can I create this new column using r and the tidyverse?

top_months_by_year <- top_months_by_year |>
    mutate(date = as.Date(paste(year, month, "01"), format = "%Y %b %d"))

approval_ratings <- approval_ratings |>
  mutate(date = as.Date(paste(year, month, "01"), format = "%Y %b %d"))

visits_by_month_ratings <- top_months_by_year |> 
  left_join(approval_ratings, by = "date") |>
  rename(year = year.x, month = month.x) |>
  select(-year.y, -month.y) |>
  na.omit(visits_by_month_ratings)
```


```{r}
#note - chatgpt wrote this code for me, I wanted to put these on the same graph but couldn't figure out the scale part
library(ggplot2)
ggplot(visits_by_month_ratings, aes(x = date)) +
  # Bar chart for total_visits
  geom_bar(aes(y = total_visits), stat = "identity", fill = "darkgray") +
  # Line graphs for approval_ratings
  geom_line(aes(y = dem_approval_pct * max(total_visits) / max(dem_approval_pct)), color = "blue") +
  geom_line(aes(y = ind_approval_pct * max(total_visits) / max(dem_approval_pct)), color = "darkgreen") +
  geom_line(aes(y = rep_approval_pct * max(total_visits) / max(dem_approval_pct)), color = "red") +
  # Y-axis scale for total_visits
  scale_y_continuous(name = "Total Visits", breaks = seq(0, 70000, by = 10000)) +
  # Secondary Y-axis scale for approval_rating
  scale_y_continuous(
    name = "Total Visits",
    sec.axis = sec_axis(~ . * max(visits_by_month_ratings$dem_approval_pct) / max(visits_by_month_ratings$total_visits), name = "Dem Approval Rating")
  ) +
  theme_gray()

#note - not sure if this is actually a valuable graph. Should we just throw out 2021 data because it's so sparse compared to 2022/23? Does this highlight fluctuations in visits or just a lack of data for earlier months? Also it wouldn't graph Dec 2022 which was the second highest month because of NA values in the approval ratings. @DEREK is there a way around this?
```
Derek's feedback:
Doesn't seem like there's any correlation there, but if you wanted to find out you could calculate a correlation coefficient or do a t-test (https://thescoop.org/data_journalism_book/basicstats.html)

Ok - let's calculate the correlation coefficient to see if these are related. I think we should just look at data from 2022 and 2023 because the 2021 data is not very robust.

```{r}
#let's take out 2021 data
top_months_by_year <- visits_by_month_ratings |> filter (date > "2021-12-31")

#now let's run the correlation test
month_ratings_post_covid |>
  select(-year, -month, -date) |>
  correlate()

```

Yup, looks like there is not really a significant relationship between visits to the White House and Biden's approval ratings for any party. 

Let's just look at how things have changed month by month.

```{r}
library(ggplot2)
ggplot(top_months_by_year, aes(x = date)) +
  # Bar chart for total_visits
  geom_bar(aes(y = total_visits), stat = "identity", fill = "darkgray") +
  theme_gray()
```

3.  How many West Wing tours were take the day cocaine was found in the White House? How does it compare to other days?

Cocaine was found in the West Wing where tours routinely went through. When evaluating the data, the first noticeable obstacle is certain labels being "WW" and others being "WEST WING." However, when looking at the capitalized WEST WING, all the accompanied caller_name_first entries say tour, which is not true for any other WW combination. When filtering for the date and WEST WING, I arrive at 1,215 people. 

```{r}
wh_visitor_data %>%  
  filter(str_detect(appt_end_date, "2023-07-02")) %>%  
  filter(str_detect(meeting_room,"WEST WING")) %>% 
  group_by(appt_end_date) %>%  
  summarize(sum= sum(total_people))
```

On July 2, cocaine was found at the White House. 1994 people visited the White House that day, presumably most of them tourists. (Hard to imagine where all those people could have fit though.) Finding out whether this was an unusal amount of people proved to be quite a challenge thanks to what we realized was a strange quirk in the data. On some days, tens of thousands of visitors went to the White House, on others, only a few hundred. The mean number of visitors was 33,801 but that's thrown off by a couple days with an insanely high number of visitors. The median is a more reasonable, 139. 

The mean being so different than the median led us to wonder if Biden had suddenly started received way more visitors as time has went on? We made a scatterplot to see, and our answer was: No, not really. There did not seem to be any rhyme or reason to when more people came, the days with thousands of people were pretty random. 
So we dug into the dates with thousands of people allegedly visiting the White House some more. We identified something pretty strange: The people racking up thousands upon thousands of visits were not the president, were not the vice president, were not important people in the cabinet. It was a pretty random assortment of comparatively unimportant people got the same number of visitors over and over again. On 12/01/2022, the day French President Emmanuel Macron visited the White House, for example, several groups of 5,619 people allegedly toured the White House to visit Kevin Ballen, a literally-just-graduated Harvard alum who completed a bachelor's. Not Joe Biden. Similar trends could be observed with other days, with the same large number of people visiting seemingly unimportant people.


```{r}
# so 1,215 people visited the white house on that date. let's check what the average amount of people on a monday is.
wh_visitor_data |> 
  mutate(day_of_week = wday(appt_end_date)) |> 
  filter(day_of_week == 1) |> 
  filter(str_detect(meeting_room,"WEST WING")) |>
  group_by(appt_end_date) |> 
  summarize(total_people = sum(total_people)) |> 
  filter(!is.na(total_people)) |> 
  summary(round(total_people))

 # Median WW visitors on a given Monday: 1,193

wh_visitors_per_day <- wh_visitor_data |> 
  mutate(date = date(appt_end_date)) |> 
  filter(str_detect(meeting_room,"WEST WING")) |>
  group_by(date) |>
  summarize(total_people = sum(total_people)) |> 
  filter(!is.na(total_people)) |>
  summary(round(total_people))

# Median : 405.0  
#Mean : 616.4

#So it looks like Mondays are particularly busy days for West Wing tours compared to other days, but there weren't a notably high number of visitors on the day they found cocaine in the White House.
```

4.  Who has sponsored the most White House visitors?
Here are the top 10 most popular people to visit in the White House:

Visitors Office - 462398	- I think this is actually meant to be Office Visitors. More on this later.
Joe	Biden	77197	- Not shocking.

Ed	Teleky	17354	- Air Force Chief Master Sgt. Edward Teleky, Air Force Band's drum major - why is he getting so many visits? He has a lot of people visiting him over and over - between 30 and 50 times.

Gionelly	Mills	12383		- Staff Assistant for the White House Visitors Office - assuming this is people going to secure some kind of visitor status/fill out paperwork etc. This checks out, the most any single person visited her was three times. They're probably stopping in at her office to get some type of credentials.

Lydia	Hecmanczuk	6956	- IT Specialist at Executive Office of the President - she probably handles when they have security IT specialists? Similar to Gionelly Mills, she has a lot of people visit her once or twice. I wonder if she has to do something to people's devices to facilitate some kind of work being done there?

Amanda	Trocola	6356 - Deputy Social Secretary to the Vice President - would be interested to see who is visiting her and why. Her top visitors are (I believe) a DC real estate agent, a florist, an NYC pianist, I think it's mostly people related to event planning. Some I couldn't pinpoint on Google though.

Dan	Via	5841 - Not sure if I found the right person but I think he's Deputy Director, Ceremonies and Protocol for the Air Force - also seems kind of weird but maybe handling events? Worth looking into further. Looks like he's getting a lot of visits from a violinist for the marines, another music person in the marines band and the CEO of a screen printing company in the area. Doesn't seem super interesting.

Kevin	Ballen	4903 - Advisor to Director of National Economic Council, previously Staff Assistant for Operations
Could be worth looking into

Peyton	Schwartz 4533 - Special Assistant to the Social Secretary & Deputy Director of Arrangements
Could be worth looking into

Kamala	Harris	4319 - Vice President
Really surprised she gets so few visits?


```{r}
wh_visitor_data |> 
  group_by(visitee_namefirst_clean, visitee_namelast_clean) |>
  summarise(visits = n()) |>
  arrange(desc(visits))
```
Let's clean our data a little bit better. 

```{r}
#used chatgpt to do this it took forever. It kept giving me code to change the last name from potus to Biden but tolower(visitee_namelast_clean) %in% c("potus") ~ "Joe" would not work. I asked it: can you help me write a case when for the following conditions: if visitee_namelast_clean == Potus or potus or POTUS,  change visitee_namelast_clean to Biden and visitee_namefirst_clean to Joe. 

#It gave me the following code but with tolower(visitee_namelast_clean) %in% c("potus") ~ "Joe" - i finally figured out it was because I had already changed the visitee_namelast_clean column in the code above so it wouldn't recognize "potus" in the visitee_namelast_clean col to change the first name to Joe. 

wh_visitor_data <- wh_visitor_data |>
  mutate(
    visitee_namelast_clean = case_when(
      tolower(visitee_namelast_clean) %in% c("potus") ~ "Biden",
      tolower(visitee_namelast_clean) %in% c("vpotus") ~ "Harris",
      tolower(visitee_namelast_clean) %in% c("flotus") ~ "Biden",
      TRUE ~ visitee_namelast_clean
    ),
    visitee_namefirst_clean = case_when(
      tolower(visitee_namelast) %in% c("potus") ~ "Joe",
      tolower(visitee_namelast) %in% c("vpotus") ~ "Kamala",
      tolower(visitee_namelast) %in% c("flotus") ~ "Jill",
      TRUE ~ visitee_namefirst_clean
    )
  )
```


```{r}
#now let's run that data again
wh_visitor_data |> 
  group_by(visitee_namefirst_clean, visitee_namelast_clean) |>
  summarise(visits = n()) |>
  arrange(desc(visits))
```
```{r}
wh_visitor_data |> 
 filter(visitee_namelast_clean == "Mills" & visitee_namefirst_clean == "Gionelly") |>
  group_by(namelast, namefirst, namemid) |>
  summarise(total_visits = n()) |>
  arrange(desc(total_visits))

wh_visitor_data |> 
 filter(visitee_namelast_clean == "Hecmanczuk" & visitee_namefirst_clean == "Lydia") |>
  group_by(namelast, namefirst, namemid) |>
  summarise(total_visits = n()) |>
  arrange(desc(total_visits))

wh_visitor_data |> 
 filter(visitee_namelast_clean == "Trocola" & visitee_namefirst_clean == "Amanda") |>
  group_by(namelast, namefirst, namemid) |>
  summarise(total_visits = n()) |>
  arrange(desc(total_visits))

wh_visitor_data |> 
 filter(visitee_namelast_clean == "Via" & visitee_namefirst_clean == "Dan") |>
  group_by(namelast, namefirst, namemid) |>
  summarise(total_visits = n()) |>
  arrange(desc(total_visits))

```


5.  What percent of visitees have just been labeled "office visitors"?
55%, roughly. That means in more than half of the visits to the White House recorded, we have no idea who the people visited. And that's only for the visits that _were_ recorded. There are all kinds of loopholes by which the White House does not have to report who's visited. But it's also possible that these 55% represent all those times, the times somebody visited and the White House doesn't want to say who. Either way, us not knowing who more than half of visitors visited is pretty sobering. 

```{r}
num_office_visitors <- wh_visitor_data %>%
  filter(str_detect(visitee_namelast, regex("office", ignore_case = TRUE))) |> 
  filter(str_detect(visitee_namefirst, regex("visitors", ignore_case = TRUE))) |> 
  summarize(count = n())  
as.character(num_office_visitors)
# result - 462349 entries with just that office visitors bs
pct_office_visitors <- num_office_visitors/847371
as.character(pct_office_visitors)
percent_value <- paste0(floor(pct_office_visitors * 100), "%")
percent_value
# So roughly 55% of all visits logged went to "office visitors."
```

6.  How often are specific people visiting? AI people? Things they want to regulate? Pick who we want to look at.

Let's see if we can find whether OpenAI lobbyists visited the president. According to Politico, there are the handful of lobbyists the company has:- Tony Samp- Chan ParkThe Politico story says that those two only started as OpenAI lobbyists in October so that doesn't help us.Let's try looking for the CEO, Samuel H Altman.

```{r}
wh_visitor_data |> 
  filter(
    str_detect(namelast, regex("altman", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^sam", ignore_case=TRUE))
         )
```

That shows us Samuel Altman. Makes sense. But what escapes me is that there is a Samuel N Altman and a Samuel H Altman. Theoretically, these should be two different people but when I search "samuel n altman -openai" with google, I get an Ancestry.com record for someone who was counted in the 1940 census, Samuel Altman the shift leader at Wallgreens, Samuel Altman the landscaper, Samuel Altman the president of a bio-nutrition labratory and Samuel Altman the lawyer in Georgia. None of those seem likely to visit the White House so it might just be a typographical error. So I'm going to assume they're all good ol' Open AI Sam.

We will take this to look at specific lobbyists who are contracted by may have visited the White House. We did this by going through the lobbying disclosures from the US House of Representatives and cross-referencing names. We searched specifically for artificial intellignece lobbyists over the past three years. 


First we will look at Julie Babayan who, according to her Linkedin, is the head of policy development at Adobe. 

In July Babayan was at the White House on the 11th and 12th. The Biden-Harris administration announced a task force related to managing the risk and "harnessing the benefits" of AI nine days after these visits ended. While Adobe did not commit until September in the second round of commitments, that could be what this was used for. 

She visited four times, all in 2023

(https://www.whitehouse.gov/briefing-room/statements-releases/2023/09/12/fact-sheet-biden-harris-administration-secures-voluntary-commitments-from-eight-additional-artificial-intelligence-companies-to-manage-the-risks-posed-by-ai/) 
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("Babayan", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^Julie", ignore_case=TRUE))
         )
```


Next we will look at Keenan Austin Reed, the executive vice president of the Alpine Group, a lobbying organization contracted by Facebook. While he's not listed as Meta's lobbyist, he does work for Lyft and ArgoAI, who specializes in driving technology. We see three visits in 2022 and one visit in 2023. 
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("austinreed", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^keenan", ignore_case=TRUE))
  )
```


Next we will look at Apple, who has a team of lobbyists on their payroll. I checked each of the following lobbyists using OpenSecrets Lobbying Profile feature and found they all exclusively lobby for Apple. 

There was no record of Jeff Dobrozsi visiting the White House.
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("dobrozsi", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^jeff", ignore_case=TRUE))
  )

```

I found multiple Robert Harris' on here, however could not find a middle intial online or anyway to identify the Robert Harris from Apple to the average person who happened to be named Robert Harris. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("harris", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^robert", ignore_case=TRUE))
  )
```


There are four visits with an April Jones, three with the middle initial of N however. All three of April N. Jones' visits came in 2023, with one being set up by Samuel Negatu on January 19, who is part of the office of US Trade Representatives. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("jones", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^april", ignore_case=TRUE))
  )
```


```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("jones", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^april", ignore_case=TRUE))
  )
```

The next lobbyist is Alexis Marks Mosher. While Alexis Marks Mosher found nothing, Alexis Marks, with the middle initial of S has made seven visits since Biden's innauguration, although two are in the same day. Once in 2021, three in 2022 and two in 2023. I found little evidence of a different Alexis Marks being relevant enough to had made so many visits. Among Marks visits was a meeting with Joe Biden. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("marks", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^alexis", ignore_case=TRUE))
  )
```

There is no evidence of Conor McGrath entering the White House. 
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("mcgrath", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^conor", ignore_case=TRUE))
  )
```

A Walter Kuhn visited twice in 2022. However, he left Apple for a position with Shopify in January of that year, both of these visits occured after, leading me to believe this isn't the Apple lobbyist visitng. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("kuhn", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^walter", ignore_case=TRUE))
  )
```

Timothy Powderly visited twice in 2021, once in 2022 and another time in 2023. 
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("powderly", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^timothy", ignore_case=TRUE))
  )
```


The next group we will be looking at are Stoney Burke and Ryan Dattlio, who work for Aquia Group LLC and are the lobbyists for Amazon Web Services. 

There is no record of Burke or Dattlio. Neither of the Amazon Web Services lobbyists 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("burke", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^stoney", ignore_case=TRUE))
  )
```


```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("Dattlio", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^Ryan", ignore_case=TRUE))
  )
```


The next name is another Amazon name. Taunja Berquam, who works for Team Hallahan and on behalf of Amazon. 

Berquam visited the White House one time in 2023. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("berquam", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^taunja", ignore_case=TRUE))
  )
```

The next name will Anthony Bedell and Clarence Williams, who represents CrowdAI. Bedell visited one in 2023. 

There is a Clarence D. Williams who visited twice, one of them being Williams III. Clarence D. Williams III is a retired partner of Shabazz Law Firm who had a 44 year law enforcement career. The lobbyist has the middle initial of L. Both have the middle intial of D, assuming they are the same person. Even if that is wrong, there is another lawyer in DC named Clarence Williams who advocates to congress. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("bedell", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^anthony", ignore_case=TRUE))
  )
```


```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("williams", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^clarence", ignore_case=TRUE))
  )
```


The next name is Robert Babcock, who is a representative of Google. There is no record of Robert Babcock visiting the White House. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("babcock", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^robert", ignore_case=TRUE))
  )
```

A Cory Crowley visited once in 2022, can't find any evidence online verifying if this is the correct one. 
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("crowley", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^cory", ignore_case=TRUE))
  )
```

The next person I'm looking at is Howard Berman, who also works with Microsoft. There is no record of a Howard Berman visiting the White House. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("berman", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^howard", ignore_case=TRUE))
  )
```

David Lugar, who works for 535 Group and represents Meta is the next name on this list. There is no record of him visiting the White House. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("lugar", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^david", ignore_case=TRUE))
  )
```

The last two names we will be looking at are Chan Park, Steve Phillips and Tony Samp, who do work for OpenAI. 
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("park", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^chan", ignore_case=TRUE))
  )
```

This gave me a number of irrelevant results, I asked ChatGPT how to only get names that are "Chan" and "Park" excluding all other names. While five visits are listed for Chan Park, four occur on April 26, which coincides with the president of the Republic of Korea visiting the White House and there being a dinner. Among the topics discussed between the two leaders was cooperation on artificial intelligence initiatives. 

Both of Park's visits were in 2023, the first was on April 17. 

```{r}
wh_visitor_data %>%
  filter(
    str_detect(namelast, regex("^PARK$", ignore_case = TRUE)) &
    str_detect(namefirst, regex("^CHAN$", ignore_case = TRUE))
  )
```

There are four visits from Steven Phillips' during the Biden presidency. However, according to Open Secrets, Steven Phillips of OpenAI has the middle intial R, which none of these have. Therefore we're not comfortable using them for data analysis. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("phillips", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^steve", ignore_case=TRUE))
  )
```


There is no record of Tony Samp having visited the White House. 
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("samp", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^tony", ignore_case=TRUE))
  )
```

In a Google Sheet I kept track of White House Visits as I went, along. Among Artifical Intelligence Lobbyists aligned with major companies, there has been a consistent growth in White House visits. This analysis measured three in 2021, seven in 2022 and 15 in 2023. 

https://docs.google.com/spreadsheets/d/1_W5cYKT39epYpnrtlrbFVyvenZn335Umg3WhmThKwPE/edit?usp=sharing
