---
title: "R Notebook"
authors: Taylor, Menna, Shaun, Joel
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
```


```{r}
## set working directory
## setwd("~/Google Drive/My Drive/JOUR472/data_journalism_2023_fall/white_house_visitor_logs")

wh_visitor_data <- read_csv("data/combined.csv")
wh_visitor_data <- wh_visitor_data |> 
  mutate(as_datetime(mdy_hm(APPT_MADE_DATE))) |>  ## when I run this, R works but tells me that "14 failed to parse." I cannot for the life of me figure out what it's unhappy about 
  mutate(as_datetime(mdy_hm(TOA))) |> 
  mutate(as_datetime(mdy_hm(TOD))) |>
  mutate(as_datetime(mdy_hm(APPT_START_DATE))) |>
  mutate(as_datetime(mdy_hm(APPT_END_DATE))) |>
  mutate(as_datetime(mdy_hm(APPT_CANCEL_DATE))) |>
  mutate(as_datetime(mdy_hm(LASTENTRYDATE))) |>
  mutate(as_datetime(mdy(RELEASEDATE))) |> 
## I'm sure there's a way to get it to not put it in a new column but I can't figure out how - I know mutate means it will convert into a new column but...it doesn't work without the "mutate." I dunno.
  clean_names()

```
Let's see when the most popular day for visiting was. To do that, first we're going to create new columns with only the dates.
```{r}
wh_visitor_data <- wh_visitor_data |> 
  mutate(as_datetime(mdy_hm(APPT_MADE_DATE))) |>  ## when I run this, R works but tells me that "14 failed to parse." I cannot for the life of me figure out what it's unhappy about 
  mutate(as_datetime(mdy(TOA))) |> 
  mutate(as_datetime(mdy(TOD))) |>
  mutate(as_datetime(mdy(APPT_START_DATE))) |>
  mutate(as_datetime(mdy(APPT_END_DATE))) |>
  mutate(as_datetime(mdy(APPT_CANCEL_DATE))) |>
  mutate(as_datetime(mdy(LASTENTRYDATE))) 

# I'm running into an error here, I'm not quite sure how to fix it - it does not (like it did before) make new columns as I've asked it to, just this time without the time. According to the Lubridate cheatsheat, my code should work so I dunno (as usual). I tried telling it "as_date" but that threw the following error:
# Error in mutate(wh_visitor_data, as_date(mdy_hm(APPT_MADE_DATE))) : 
# Caused by error in `h()`:
# ! error in evaluating the argument 'x' in selecting a method for function 'as_date': object 'APPT_MADE_DATE' not found
# Whenever I've fixed the above problems, here's what I'd do next:

wh_visitor_data
  group_by(date_of_visit) |> 
  summarize(count=n())
  
# once we've figured out the above headaches that should work fine.
```
Let's see if we can find whether OpenAI lobbyists visited the president.
According to Politico, here are the handful of lobbyists the company has:
-   Tony Samp
-   Chan Park
The Politico story says that those two only started as OpenAI lobbyists in October so that doesn't help us.
Let's try looking for the CEO, Samuel H Altman.
```{r}
clean_wh_visitor_data |> 
  filter(
    str_detect(namelast, regex("altman", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^sam", ignore_case=TRUE))
         )
```
That shows us Samuel Altman.
Makes sense.
But what escapes me is that there is a Samuel N Altman and a Samuel H Altman.
Theoretically, these should be two different people but when I search "samuel n altman -openai" with google, I get an Ancestry.com record for someone who was counted in the 1940 census, Samuel Altman the shift leader at Wallgreens, Samuel Altman the landscaper, Samuel Altman the president of a bio-nutrition labratory and Samuel Altman the lawyer in Georgia.
None of those seem likely to visit the White House so it might just be a typographical error.
Anyhow, assuming all of them are Samuel Altman, CEO of OpenAI:

What are the most popular months and dates? What political events correlate with those?

{r}
#top 5 dates for visitors


Has the amount of visitors fluctuated during Biden presidency (see if this correlates with rise in drop of approval rating)

```{r}

```


West Wing Tours are selective, how many are being taken?

```{r}

```


Who has sponsored the most White House visitors?

```{r}

```


What percent of visitors have just been labeled "office visitors"?

```{r}
clean_wh_visitor_data %>%
  filter(str_detect(visitee_namelast, "Office")) |> 
  filter(str_detect(visitee_namefirst, "Visitors"))

clean_wh_visitor_data |> 
  group_by(visitee_namelast, visitee_namefirst) |> 
  summarize(total_visits = n()) |> 
  arrange(desc(total_visits))

clean_wh_visitor_data |> 
  group_by(vistitee_namefirst) 
```



How often are specific people visiting? AI people? Things they want to regulate? Pick who we want to look at.

```{r}

```




