---
title: "R Notebook"
authors: Taylor, Menna, Shaun, Joel
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
```


```{r}
## set working directory
## setwd("~/Google Drive/My Drive/JOUR472/data_journalism_2023_fall/white_house_visitor_logs")

wh_visitor_data <- read_csv("data/combined.csv")
wh_visitor_data <- wh_visitor_data |> 
  mutate(as_datetime(mdy_hm(APPT_END_DATE))) |>   ## when I run this, R works but tells me that "14 failed to parse." I cannot for the life of me figure out what it's unhappy about 
# mutate(as_datetime(mdy_hm(TOA))) |> 
#  mutate(as_datetime(mdy_hm(TOD))) |>
#  mutate(as_datetime(mdy_hm(APPT_START_DATE))) |>
#  mutate(as_datetime(mdy_hm(APPT_END_DATE))) |>
#  mutate(as_datetime(mdy_hm(APPT_CANCEL_DATE))) |>
# mutate(as_datetime(mdy_hm(LASTENTRYDATE))) |>
# mutate(as_datetime(mdy(RELEASEDATE))) |> 
## I'm sure there's a way to get it to not put it in a new column but I can't figure out how - I know mutate means it will convert into a new column but...it doesn't work without the "mutate." I dunno.
  clean_names()

```
```{r}
wh_visitor_data <- wh_visitor_data |> 
  mutate(as_datetime(mdy_hm(appt_end_date)))

wh_visitor_data |> 
  filter(str_detect(appt_end_date, "2023-07-02")) |> 
  group_by(appt_end_date) |> 
  summarize(sum= sum(total_people))

wh_visitor_data |> 
  filter(str_detect(appt_end_date, "2023-07-02")) |> 
  mutate(wday(appt_end_date))
# ok so we just found out in a way too complex way that July 2 was a Monday. Now let's run this code again to see what the average number of visitors on a monday is.

wh_visitor_data |> 
  mutate(day_of_week = wday(appt_end_date)) |> 
  filter(day_of_week == 1) |> 
  group_by(appt_end_date) |> 
  summarize(count = n()) |> 
  summarize(mean(count))

# ok so the mean is 299.627 entries per day. but i realize that won't work because each row says how many people visited on that one day and that number was not just 1. let's try about it again
wh_visitor_data |> 
  

```
So 324 people visited the white house on cocaine day.

Let's see when the most popular day for visiting was. To do that, first we're going to create new columns with only the dates.
```{r}
wh_visitor_data <- wh_visitor_data |> 
  mutate(as_datetime(mdy_hm(APPT_MADE_DATE))) |>  ## when I run this, R works but tells me that "14 failed to parse." I cannot for the life of me figure out what it's unhappy about 
  mutate(as_datetime(mdy(TOA))) |> 
  mutate(as_datetime(mdy(TOD))) |>
  mutate(as_datetime(mdy(APPT_START_DATE))) |>
  mutate(as_datetime(mdy(APPT_END_DATE))) |>
  mutate(as_datetime(mdy(APPT_CANCEL_DATE))) |>
  mutate(as_datetime(mdy(LASTENTRYDATE))) 

# I'm running into an error here, I'm not quite sure how to fix it - it does not (like it did before) make new columns as I've asked it to, just this time without the time. According to the Lubridate cheatsheat, my code should work so I dunno (as usual). I tried telling it "as_date" but that threw the following error:
# Error in mutate(wh_visitor_data, as_date(mdy_hm(APPT_MADE_DATE))) : 
# Caused by error in `h()`:
# ! error in evaluating the argument 'x' in selecting a method for function 'as_date': object 'APPT_MADE_DATE' not found
# Whenever I've fixed the above problems, here's what I'd do next:

wh_visitor_data |> 
  group_by(date_of_visit) |> 
  summarize(count=n())
  
# once we've figured out the above headaches that should work fine.
```
Let's see if we can find whether OpenAI lobbyists visited the president.
According to Politico, here are the handful of lobbyists the company has:
-   Tony Samp
-   Chan Park
The Politico story says that those two only started as OpenAI lobbyists in October so that doesn't help us.
Let's try looking for the CEO, Samuel H Altman.
```{r}
clean_wh_visitor_data |> 
  filter(
    str_detect(namelast, regex("altman", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^sam", ignore_case=TRUE))
         )
```
That shows us Samuel Altman.
Makes sense.
But what escapes me is that there is a Samuel N Altman and a Samuel H Altman.
Theoretically, these should be two different people but when I search "samuel n altman -openai" with google, I get an Ancestry.com record for someone who was counted in the 1940 census, Samuel Altman the shift leader at Wallgreens, Samuel Altman the landscaper, Samuel Altman the president of a bio-nutrition labratory and Samuel Altman the lawyer in Georgia.
None of those seem likely to visit the White House so it might just be a typographical error.
Anyhow, assuming all of them are Samuel Altman, CEO of OpenAI:

What are the most popular months and dates? What political events correlate with those?

{r}
#top 5 dates for visitors


Has the amount of visitors fluctuated during Biden presidency (see if this correlates with rise in drop of approval rating)

```{r}

```


West Wing Tours are selective, how many are being taken?

```{r}

```


Who has sponsored the most White House visitors?

```{r}

```


What percent of visitors have just been labeled "office visitors"?

```{r}
wh_visitor_data %>%
  filter(str_detect(visitee_namelast, regex("office", ignore_case = TRUE))) |> 
  filter(str_detect(visitee_namefirst, regex("visitors", ignore_case = TRUE))) |> 
  summarize(count = n())  
# result - 462349 entries with just that office visitors bs
  
462349/847371
```

```{sql connection=}
SELECT *
FROM clean_wh_visitor_data
```


How often are specific people visiting? AI people? Things they want to regulate? Pick who we want to look at.

```{r}

```




