---
title: "R Notebook"
output: html_notebook
---

---
title: "R Notebook"
authors: Taylor, Menna, Shaun, Joel
output: html_notebook
---

Memo link: https://docs.google.com/document/d/1j_19XNibAk0Modj5MFnaPfIz1M4JZ1dzW1s5KP2PIQM/edit?usp=sharing 

```{r}
library(tidyverse)
library(tools)
library(janitor)
library(lubridate)
library(ggplot2) #make sure you have ggplot 2 installed, not ggplot
options(scipen=999)

```


```{r}
install.packages("corrr")
```

```{r}
###asked chatgpt how to turn a character column into a date time column, gave it some examples from our code and int told me to do this:
wh_visitor_data <- read_csv("data/cleaned_OPENREFINE.csv") %>% 
  clean_names()
wh_visitor_data <- wh_visitor_data %>% 
  mutate(appt_end_date = mdy_hm(appt_end_date))
  
```

Let's take a quick look at the data now that we've changed these columns into dates.

```{r}
glimpse(wh_visitor_data)
```

Here are our new questions we'd like to answer: What are the most popular times / dates / months for White House visitors and how do they correlate with important political events? Has the amount of visitors fluctuated during Biden presidency (see if this correlates with rise in drop of approval rating) West Wing Tours are selective, how many are being taken? Who has sponsored the most White House visitors - e.g. who is getting the most visitors? What percent of visitors have just been labeled "office visitors"? If it's a significant percentage, that would seem to defeat the point of releasing these logs. Not much better than the Trump administration.

Immediate limitations: 1. There are a lot of NAs in here where dates should go. I think the most reliable date is the appointment end date, which means the times are probably useless for us. 2. Dates and times aren't always written the same way. Is that going to cause us problems? We should probably check with Derek about that. 3. We wanted to look at how many folks are labeled office visitor - we're gonna have to find all instances of that, they're not marked the same. 4. We had a question about west wing visits. I don't see any identifying data here to show which ones are west wing visits, although there are a lot of confusing column names and abbreviations. 5. There are a ot of confusing column names and abbreviations - we will need to dig deeper into that to find out what we can actually look at here. 6. I still really want to know what types of people are visiting. Could we at least match congress people up using the propublica dataset and see how many folks are not house or senate members? We may be able to find a dataset of foreign powers we can match up as well.

Let's see when the most popular day for visiting was. To do that, first we're going to create new columns with only the dates.

```{r}
####mutate(toa_time = ymd_hms(toa) %>%  time()) we need to figure out how to separate out time from these two columns as well

wh_visitor_data %>% 
  group_by(toa) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
  
# note - since there are so many NA dates here, I think we should group by appt date.

wh_visitors_by_date <- wh_visitor_data %>% 
  group_by(appt_start_date) %>% 
  summarize(total_visits = n()) %>% 
  arrange(desc(total_visits))

### just kidding, there are a ton of NA dates still. maybe appt end date?

wh_visitors_by_end_date <- wh_visitor_data %>% 
  group_by(appt_end_date) %>% 
  summarize(total_visits = n()) %>% 
  arrange(desc(total_visits))

### this seems better
```


```{r}
###derek asked if the NA dates were specific to certain offices. how do we figure that out?

wh_visitor_data %>%  filter(if_any(-toa, is.na)) %>% 
  group_by(meeting_loc) %>% 
  summarise(count = n())
```

Question 1.  What are the most popular months and dates? What political events correlate with those?

WHERE WE ARE AT RIGHT NOW: 
Need to do some digging into the names that appear on these dates per Derek’s feedback. I think we should look at the dates when Menna filtered for groups smaller than 100. I think it would be worth looking at either the top 10 or looking at top dates filtering for groups smaller than 500 or something to se how it varies. We should def look at the data for the mysterious 3/18/23 date.

```{r}

###mutate(toa_time = ymd_hms(toa) %>%  time()) we need to figure out how to separate out time from these two columns as well

wh_visitor_data %>% 
  group_by(toa) %>%  
  summarize(count=n()) %>% 
  arrange(desc(count))
  
# note - since there are so many NA dates here, I think we should group by appt date.

wh_visitors_by_date <- wh_visitor_data %>% 
  group_by(appt_start_date) %>% 
  summarize(total_visits = n()) %>% 
  arrange(desc(total_visits))
%>% %>% 
### just kidding, there are a ton of NA dates still. maybe appt end date?

wh_visitors_by_end_date <- wh_visitor_data  %>% 
  group_by(appt_end_date) %>% 
  summarize(total_visits = n()) %>% 
  arrange(desc(total_visits))

#That looks better. Let's use appt end date because it seems to be the date col with the most data and fewest NAs.

wh_visitors_by_end_date

```

Top 5 Dates: 

1. 06-22-2023: President Biden and Prime Minister Modi of the Republic of India in Joint Press Conference
2. 12-23-2022: President Biden passes a successful bipartisan funding bill
3. 04-26-2023:President Biden and President Yoon Suk Yeol of the Republic of Korea to commemorate the 70th anniversary of their alliance
4. 07-04-2023: Independence Day event at the White House hosted by President Biden and First Lady Jill Biden (recurring date with a big total of visitors)
5. 12-13-2022: President Biden signs the Respect for Marriage Act, legalizing gay marriage

Another important date: July 2, the day cocaine was found in the White House.

This is actually pretty interesting! It's somewhat surprising that the top date doesn't solely have to do with the United States itself, but I remember driving through Capitol Hill several times that week and seeing the huge impact (both good and bad) that Modi's visit had on the city. The majority of these highest visited dates correlate with important international visits, United States-specific holidays, and the passage of significant things into law.

We think it may actually be more interesting to look at data where groups were less than 100 because those may be attributed to big events or holidays, which we could have guessed. What are the most popular dates and months for regular folks?

```{r}
wh_visitor_data %>% 
  group_by(appt_end_date) %>%  
  filter(total_people <100) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
```

Top 5 Dates: 

1. 12-23-2022: President Biden passes a successful bipartisan funding bill
2. 03-18-2023: literally NO idea nothing is showing up (pls help)
3. 05-11-2023: This was a big day at The House because the Biden-Harris Administration made a lot of big decisions about Global Health, Environmental Protection and National Conservation, Immigration, and International/National Emergency Funding
4. 03-25-2023: President Biden and President Alberto Fernandez of Argentina 
5. 05-12-2023: Continuation of the decisions made on 5/11, as well as the start of Infrastructure Week.

This is so interesting because, aside from 12/23, none of the dates repeat in the data without big organized groups. This shows that big organized groups frequently visit on holidays and days with big international repercussions, while more national-focused events and big decisions are primarily attended by individuals or folks in smaller groups.

Note from Taylor - Can we look at the data for 3/18/23 to see who visited? Maybe it had something to do with St. Patrick's Day? Biden is Irish Catholic. Not sure if people actually do anything for that though.

2.  Has the amount of visitors fluctuated during Biden presidency (see if this correlates with rise in drop of approval rating)

ANSWER: Visits in 2021 were super minimal. In 2022, visits were fairly 

```{r}
### Note - asked ChatGPT how to make the month column have labels and gave it my mutate code creating a new col with lubridate
#reading in approval ratings data from gallup
approval_ratings <- read_csv("data/biden_approval_gallup.csv", col_names = FALSE)

colnames(approval_ratings) <- c("poll_date", "rep_approval_pct", "ind_approval_pct", "dem_approval_pct")
  
approval_ratings <- approval_ratings %>% 
  separate(poll_date, into = c("year", "month", sep = "-")) %>% 
  select(year, month, rep_approval_pct, ind_approval_pct, dem_approval_pct)

approval_ratings <- approval_ratings %>% 
  mutate(date = paste(year, month, sep = " "))

#making a df with top months for visitors
wh_visitor_by_month <- wh_visitors_by_end_date %>%  
 mutate(date_time = ymd_hms(appt_end_date),
  month = month.abb[month(appt_end_date)])

wh_visitor_by_month <- wh_visitor_by_month %>% 
  group_by(month) %>% 
  summarise(total_visits = sum(total_visits))

#last one not really helpful bc didn't sort by year. Now doing the same with year and month
wh_visitors_month_year <- wh_visitors_by_end_date %>%  
 mutate(date_time = ymd_hms(appt_end_date),
  month = month.abb[month(appt_end_date)],
  year = year(date_time))

top_months_by_year <- wh_visitors_month_year %>% 
  group_by(year, month) %>% 
   summarise(total_visits = sum(total_visits))

```



```{r}
#not sure if that was useful. I think I need to join these dataframes. Asked chatgpt for help with this because I needed to join on the date column and didn't have one with year and month in it for top_months_by_year. Asked it: I want to create a date column that looks like this: 2022 Jan, 2022 Feb, 2022 March. Currently I have a year column that looks like this: 2022, 2022, 2022, and a month column that looks like this: Jan, Feb, March. How can I create this new column using r and the tidyverse?

top_months_by_year <- top_months_by_year %>% 
    mutate(date = as.Date(paste(year, month, "01"), format = "%Y %b %d"))

approval_ratings <- approval_ratings %>% 
  mutate(date = as.Date(paste(year, month, "01"), format = "%Y %b %d"))

visits_by_month_ratings <- top_months_by_year %>%  
  left_join(approval_ratings, by = "date") %>% 
  rename(year = year.x, month = month.x) %>% 
  select(-year.y, -month.y) %>% 
  na.omit(visits_by_month_ratings)
```


```{r}
#note - chatgpt wrote this code for me, I wanted to put these on the same graph but couldn't figure out the scale part
library(ggplot2)
ggplot(visits_by_month_ratings, aes(x = date)) +
  # Bar chart for total_visits
  geom_bar(aes(y = total_visits), stat = "identity", fill = "darkgray") +
  # Line graphs for approval_ratings
  geom_line(aes(y = dem_approval_pct * max(total_visits) / max(dem_approval_pct)), color = "blue") +
  geom_line(aes(y = ind_approval_pct * max(total_visits) / max(dem_approval_pct)), color = "darkgreen") +
  geom_line(aes(y = rep_approval_pct * max(total_visits) / max(dem_approval_pct)), color = "red") +
  # Y-axis scale for total_visits
  scale_y_continuous(name = "Total Visits", breaks = seq(0, 70000, by = 10000)) +
  # Secondary Y-axis scale for approval_rating
  scale_y_continuous(
    name = "Total Visits",
    sec.axis = sec_axis(~ . * max(visits_by_month_ratings$dem_approval_pct) / max(visits_by_month_ratings$total_visits), name = "Dem Approval Rating")
  ) +
  theme_gray()

#note - not sure if this is actually a valuable graph. Should we just throw out 2021 data because it's so sparse compared to 2022/23? Does this highlight fluctuations in visits or just a lack of data for earlier months? Also it wouldn't graph Dec 2022 which was the second highest month because of NA values in the approval ratings. @DEREK is there a way around this?
```
Derek's feedback:
Doesn't seem like there's any correlation there, but if you wanted to find out you could calculate a correlation coefficient or do a t-test (https://thescoop.org/data_journalism_book/basicstats.html)

Ok - let's calculate the correlation coefficient to see if these are related. I think we should just look at data from 2022 and 2023 because the 2021 data is not very robust.

```{r}
#let's take out 2021 data
month_ratings_post_covid <- visits_by_month_ratings %>%  filter (date > "2021-12-31")

#now let's run the correlation test
month_ratings_post_covid %>% 
  select(-year, -month, -date) %>% 
  #correlate()
```

Yup, looks like there is not really a significant relationship between visits to the White House and Biden's approval ratings for any party. 

Let's just look at how things have changed month by month in 2022 or 2023. 

```{r}
library(ggplot2)
ggplot(month_ratings_post_covid, aes(x = date)) +
  # Bar chart for total_visits
  geom_bar(aes(y = total_visits), stat = "identity", fill = "darkgray") +
  theme_gray()
```

3.  West Wing Tours are selective, how many are being taken?

WHERE WE ARE AT RIGHT NOW:
Shaun are you working on this one? There’s nothing in the group notebook for this but we do have the analysis done looking at the data for the cocaine day. Might be good to look at the # of ppl who did west wing tours that day specifically and the mean / median # of ppl who visit for west wing tours normally.

```{r}
# July 2 calculation
wh_visitor_data %>%  
  filter(str_detect(appt_end_date, "2023-07-02")) %>%  
  group_by(appt_end_date) %>%  
  summarize(sum= sum(total_people))
```

Let's look at the day cocaine was found in the White House. The Cocaine was found in the West Wing where tours routinely went through. When evaluating the data, the first noticeable obstacle is certain labels being "WW" and others being "WEST WING." However, when looking at the capitalized WEST WING, all the accompanied caller_name_first entries say tour, which is not true for any other WW combination. When filtering for the date and WEST WING, I arrive at 1,215 people. 

```{r}
wh_visitor_data %>%  
  filter(str_detect(appt_end_date, "2023-07-02")) %>%  
  filter(str_detect(meeting_room,"WEST WING")) %>% 
  group_by(appt_end_date) %>%  
  summarize(sum= sum(total_people))
```


```{r}
# so 1994 people visited the white house on that date. let's check what the average amount of people on monday is.
wh_visitor_data %>%  
  mutate(day_of_week = wday(appt_end_date)) %>%  
  filter(day_of_week == 1) %>%  
  group_by(appt_end_date) %>%  
  summarize(total_people = sum(total_people)) %>%  
  filter(!is.na(total_people)) %>%  
  summary(round(total_people))

wh_visitors_per_day <- wh_visitor_data %>%  
  mutate(date = date(appt_end_date)) %>%  
  group_by(date) %>%  
  summarize(total_people = sum(total_people)) %>%  
  filter(!is.na(total_people)) 

#answer - 33,801 is the average number of visitors via the mean but that's thrown off by a couple days with an insanely high number of visitors. the median is more reasonable, 139.

# funny. let's make a scatterplot with a line of best fit to see what the distribution is. perhaps the number of visitors has increased as time as went on?
# without the function (thanks bard) ggplot was displaying like 2e19 which if i really think hard i can decipher but i am lazy so i made it display in a reasonable format
format_numbers <- function(x) format(x, big.mark = ",", trim = 0)

scatter_visitors_day <- 
  ggplot(data=wh_visitors_per_day) +
  geom_point(aes(x=date, y=total_people))

scatter_visitors_day <- scatter_visitors_day + scale_y_continuous(labels = format_numbers)
```


```{r}
# It looks like the answer is not really. The days with tens of thousands of people seem to be big outliers, something is screwy with the white house's data
wh_visitors_per_day %>%  
  arrange(desc(total_people))
# on 12/01/2022, for example, several groups of 5,619 people allegedly toured the white house to visit kevin ballen. i looked him up and it looks like he's a really annoying (at least from his looks) literally-just-graduated Harvard alum who completed a bachelor's. 
# edit: the more i read about him the more i am convinced he is an annoying person. from a crimson puff piece:
# Kevin Ballen didn’t plan on taking two gap years. But he did intend to live a life less constrained by society’s expectations. [...] But a single moment made him rethink his altruistic efforts. It happened outside a local supermarket at a fundraiser for a day care that served homeless children. At the last minute, student leaders were unable to attend and Ballen was put in charge. “It was a realization of agency and what was possible for a young person like me,” he said. “I was able to lead a team, to make tangible change, to use my creativity.” 
# sounds like someone owed this kid's parent a favor. cHrist. 
# edit #2: it gets worse the more i look. this kid's thesis topic - i shit you not - that got featured on the harvard tiktok was literally "young people don't love taxes as much as he does." 

# i decided to take a look at what the actual hell is happening on one of the days with a gigantic number of people. first, Dec. 1 2022. Macron visited the White House on that day.
wh_visitor_data %>%  
     filter(str_detect(appt_end_date, "2022-12-01")) %>%  
     select(total_people, namelast, namefirst, visitee_namelast, visitee_namefirst) %>%  
     arrange(desc(total_people))

# What's really strange is that 5619 people are visiting Ballen over and over. And they're not vising Joe Biden. I think he's either running a covert drug operation, an illicit prostitution ring (maybe Pizzagate wasn't that far off after all??), or this is a weird data keeping error. Let's see if it continues with the other days.

wh_visitor_data %>%  
     filter(str_detect(appt_end_date, "2022-09-13")) %>%  
     select(total_people, namelast, namefirst, visitee_namelast, visitee_namefirst) %>%  
     arrange(desc(total_people))

# So on this day, which is the day with the second most people visiting the white house, we see the weird repetition of the number of people visiting again, 4478 this time visiting Biden. Let's look at July 4, 2022. A large amount of people would make sense on that day.

wh_visitor_data %>%  
  filter(str_detect(appt_end_date, "2023-07-04")) %>%  
  select(namefirst, namelast,  visitee_namefirst, visitee_namelast, total_people) %>%  
  arrange(desc(total_people))

# So this time the magic number is 1182. They're all (allegedly) visiting Bonnie Casillas, the associate director for White House operations. Let's see if there are any rows with a reasonable amount of people visiting.
wh_visitor_data %>%  
  filter(str_detect(appt_end_date, "2023-07-04")) %>%  
  filter(total_people != 1182) %>%  
  select(namefirst, namelast,  visitee_namefirst, visitee_namelast, total_people) %>%  
  arrange(desc(total_people))

# Uhh. Now we have 788 people visiting Biden over and over again. Also strange. Let's see what happens when we filter out both 1182 and 788 

wh_visitor_data %>%  
  filter(str_detect(appt_end_date, "2023-07-04")) %>%  
    filter(total_people != 1182 & total_people != 788) %>%  
  select(namefirst, namelast,  visitee_namefirst, visitee_namelast, total_people) %>%  
  arrange(desc(total_people))

# Now 752 is the number du jour. Jesus christ.
wh_visitor_data %>%  
  filter(str_detect(appt_end_date, "2023-07-04")) %>%  
  filter(total_people != 1182 & total_people != 788 & total_people !=752) %>%  
  select(namefirst, namelast,  visitee_namefirst, visitee_namelast, total_people) %>%  
  arrange(desc(total_people))

# This is really weird. 
wh_visitor_data %>%  
  filter(str_detect(appt_end_date, "2023-07-04")) %>%  
  unique(total_people) %>%  
  select(namefirst, namelast,  visitee_namefirst, visitee_namelast, total_people) %>%  
  arrange(desc(total_people))

wh_repeats <- table(wh_visitor_data$total_people)
wh_repeats <- as.data.frame(wh_repeats)

wh_repeats %>%
  ggplot(aes(x = Freq)) +
  geom_histogram(breaks=c(0, 10000, 20000, 30000, 40000, 50000, 60000)) 


# So we know that usually, 
```


4.  Who has sponsored the most White House visitors?
Here are the top 10 most popular people to visit in the White House:

Visitors Office - 462398	- I think this is actually meant to be Office Visitors. More on this later.
Joe	Biden	77197	- Not shocking.
Ed	Teleky	17354	- Air Force Chief Master Sgt. Edward Teleky, Air Force Band's drum major - why is he getting so many visits?
Gionelly	Mills	12383		- Staff Assistant for the White House Visitors Office
Lydia	Hecmanczuk	6956	- IT Specialist at Executive Office of the President
Amanda	Trocola	6356 - Deputy Social Secretary to the Vice President
Dan	Via	5841		
Kevin	Ballen	4903		
Peyton	Schwartz	4533		
Kamala	Harris	4319 - Vice President

WHERE WE ARE AT: 
Mostly done with this. We should do some digging into the top ppl though. I can tackle this one - need to fix Kamala Harris with case_when

This won't load properly because we have to upload the file and I missed a potus column when I was cleaning, but after office visitors, it's Biden, then Ed Teleky (17k visitors) and then Gionelly Mills (almost 13k visits).

```{r}
wh_visitor_data %>%  
  group_by(visitee_namefirst_clean, visitee_namelast_clean) %>% 
  summarise(visits = n()) %>% 
  arrange(desc(visits))
```
Let's clean our data a little bit better. 

```{r}
#used chatgpt to do this it took forever. It kept giving me code to change the last name from potus to Biden but tolower(visitee_namelast_clean) %in% c("potus") ~ "Joe" would not work. I asked it: can you help me write a case when for the following conditions: if visitee_namelast_clean == Potus or potus or POTUS,  change visitee_namelast_clean to Biden and visitee_namefirst_clean to Joe. 

#It gave me the following code but with tolower(visitee_namelast_clean) %in% c("potus") ~ "Joe" - i finally figured out it was because I had already changed the visitee_namelast_clean column in the code above so it wouldn't recognize "potus" in the visitee_namelast_clean col to change the first name to Joe. 

wh_visitor_data <- wh_visitor_data %>% 
  mutate(
    visitee_namelast_clean = case_when(
      tolower(visitee_namelast_clean) %in% c("potus") ~ "Biden",
      tolower(visitee_namelast_clean) %in% c("vpotus") ~ "Harris",
      tolower(visitee_namelast_clean) %in% c("flotus") ~ "Biden",
      TRUE ~ visitee_namelast_clean
    ),
    visitee_namefirst_clean = case_when(
      tolower(visitee_namelast) %in% c("potus") ~ "Joe",
      tolower(visitee_namelast) %in% c("vpotus") ~ "Kamala",
      tolower(visitee_namelast) %in% c("flotus") ~ "Jill",
      TRUE ~ visitee_namefirst_clean
    )
  )
```


```{r}
#now let's run that data again
wh_visitor_data %>%  
  group_by(visitee_namefirst_clean, visitee_namelast_clean) %>% 
  summarise(visits = n()) %>% 
  arrange(desc(visits))
```


5.  What percent of visitors have just been labeled "office visitors"?

WHERE WE ARE AT:
Done. Not that interesting but we need to do a little write up for it maybe we can make it sound more interesting highlight how this label is fucking up the data

```{r}
wh_visitor_data %>%
  filter(str_detect(visitee_namelast, regex("office", ignore_case = TRUE))) %>%  
  filter(str_detect(visitee_namefirst, regex("visitors", ignore_case = TRUE))) %>%  
  summarize(count = n())  
# result - 462349 entries with just that office visitors bs
  
462349/847371
```

6.  How often are specific people visiting? AI people? Things they want to regulate? Pick who we want to look at.



```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("altman", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^sam", ignore_case=TRUE))
         )
```

That shows us Samuel Altman. Makes sense. But what escapes me is that there is a Samuel N Altman and a Samuel H Altman. Theoretically, these should be two different people but when I search "samuel n altman -openai" with google, I get an Ancestry.com record for someone who was counted in the 1940 census, Samuel Altman the shift leader at Wallgreens, Samuel Altman the landscaper, Samuel Altman the president of a bio-nutrition labratory and Samuel Altman the lawyer in Georgia. None of those seem likely to visit the White House so it might just be a typographical error. So I'm going to assume they're all good ol' Open AI Sam.


We will take this to look at specific lobbyists who are contracted by may have visited the White House. We did this by going through the lobbying disclosures from the US House of Representatives and cross-referencing names. We searched specifically for artificial intellignece lobbyists over the past three years. 


First we will look at Julie Babayan who, according to her Linkedin, is the head of policy development at Adobe. 

In July Babayan was at the White House on the 11th and 12th. The Biden-Harris administration announced a task force related to managing the risk and "harnessing the benefits" of AI nine days after these visits ended. While Adobe did not commit until September in the second round of commitments, that could be what this was used for. 

She visited four times, all in 2023

(https://www.whitehouse.gov/briefing-room/statements-releases/2023/09/12/fact-sheet-biden-harris-administration-secures-voluntary-commitments-from-eight-additional-artificial-intelligence-companies-to-manage-the-risks-posed-by-ai/) 
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("Babayan", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^Julie", ignore_case=TRUE))
         )
```


Next we will look at Keenan Austin Reed, the executive vice president of the Alpine Group, a lobbying organization contracted by Facebook. While he's not listed as Meta's lobbyist, he does work for Lyft and ArgoAI, who specializes in driving technology. We see three visits in 2022 and one visit in 2023. 
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("austinreed", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^keenan", ignore_case=TRUE))
  )
```


Next we will look at Apple, who has a team of lobbyists on their payroll. I checked each of the following lobbyists using OpenSecrets Lobbying Profile feature and found they all exclusively lobby for Apple. 

There was no record of Jeff Dobrozsi visiting the White House.
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("dobrozsi", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^jeff", ignore_case=TRUE))
  )

```

I found multiple Robert Harris' on here, however could not find a middle intial online or anyway to identify the Robert Harris from Apple to the average person who happened to be named Robert Harris. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("harris", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^robert", ignore_case=TRUE))
  )
```


There are four visits with an April Jones, three with the middle initial of N however. All three of April N. Jones' visits came in 2023, with one being set up by Samuel Negatu on January 19, who is part of the office of US Trade Representatives. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("jones", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^april", ignore_case=TRUE))
  )
```


```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("jones", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^april", ignore_case=TRUE))
  )
```

The next lobbyist is Alexis Marks Mosher. While Alexis Marks Mosher found nothing, Alexis Marks, with the middle initial of S has made seven visits since Biden's innauguration, although two are in the same day. Once in 2021, three in 2022 and two in 2023. I found little evidence of a different Alexis Marks being relevant enough to had made so many visits. Among Marks visits was a meeting with Joe Biden. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("marks", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^alexis", ignore_case=TRUE))
  )
```

There is no evidence of Conor McGrath entering the White House. 
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("mcgrath", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^conor", ignore_case=TRUE))
  )
```

A Walter Kuhn visited twice in 2022. However, he left Apple for a position with Shopify in January of that year, both of these visits occured after, leading me to believe this isn't the Apple lobbyist visitng. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("kuhn", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^walter", ignore_case=TRUE))
  )
```

Timothy Powderly visited twice in 2021, once in 2022 and another time in 2023. 
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("powderly", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^timothy", ignore_case=TRUE))
  )
```


The next group we will be looking at are Stoney Burke and Ryan Dattlio, who work for Aquia Group LLC and are the lobbyists for Amazon Web Services. 

There is no record of Burke or Dattlio. Neither of the Amazon Web Services lobbyists 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("burke", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^stoney", ignore_case=TRUE))
  )
```


```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("Dattlio", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^Ryan", ignore_case=TRUE))
  )
```


The next name is another Amazon name. Taunja Berquam, who works for Team Hallahan and on behalf of Amazon. 

Berquam visited the White House one time in 2023. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("berquam", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^taunja", ignore_case=TRUE))
  )
```

The next name will Anthony Bedell and Clarence Williams, who represents CrowdAI. Bedell visited one in 2023. 

There is a Clarence D. Williams who visited twice, one of them being Williams III. Clarence D. Williams III is a retired partner of Shabazz Law Firm who had a 44 year law enforcement career. The lobbyist has the middle initial of L. Both have the middle intial of D, assuming they are the same person. Even if that is wrong, there is another lawyer in DC named Clarence Williams who advocates to congress. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("bedell", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^anthony", ignore_case=TRUE))
  )
```

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("williams", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^clarence", ignore_case=TRUE))
  )
```


The next name is Robert Babcock, who is a representative of Google. There is no record of Robert Babcock visiting the White House. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("babcock", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^robert", ignore_case=TRUE))
  )
```

A Cory Crowley visited once in 2022, can't find any evidence online verifying if this is the correct one. 
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("crowley", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^cory", ignore_case=TRUE))
  )
```

The next person I'm looking at is Howard Berman, who also works with Microsoft. There is no record of a Howard Berman visiting the White House. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("berman", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^howard", ignore_case=TRUE))
  )
```

David Lugar, who works for 535 Group and represents Meta is the next name on this list. There is no record of him visiting the White House. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("lugar", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^david", ignore_case=TRUE))
  )
```

The last two names we will be looking at are Chan Park, Steve Phillips and Tony Samp, who do work for OpenAI. 
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("park", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^chan", ignore_case=TRUE))
  )
```

This gave me a number of irrelevant results, I asked ChatGPT how to only get names that are "Chan" and "Park" excluding all other names. While five visits are listed for Chan Park, four occur on April 26, which coincides with the president of the Republic of Korea visiting the White House and there being a dinner. Among the topics discussed between the two leaders was cooperation on artificial intelligence initiatives. 

Both of Park's visits were in 2023, the first was on April 17. 

```{r}
wh_visitor_data %>%
  filter(
    str_detect(namelast, regex("^PARK$", ignore_case = TRUE)) &
    str_detect(namefirst, regex("^CHAN$", ignore_case = TRUE))
  )
```

There are four visits from Steven Phillips' during the Biden presidency. However, according to Open Secrets, Steven Phillips of OpenAI has the middle intial R, which none of these have. Therefore we're not comfortable using them for data analysis. 

```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("phillips", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^steve", ignore_case=TRUE))
  )
```



There is no record of Tony Samp having visited the White House. 
```{r}
wh_visitor_data %>%  
  filter(
    str_detect(namelast, regex("samp", ignore_case = TRUE))
    &
    str_detect(namefirst, regex("^tony", ignore_case=TRUE))
  )
```

In a Google Sheet I kept track of White House Visits as I went, along. Among Artifical Intelligence Lobbyists aligned with major companies, there has been a consistent growth in White House visits. This analysis measured three in 2021, seven in 2022 and 15 in 2023. 

https://docs.google.com/spreadsheets/d/1_W5cYKT39epYpnrtlrbFVyvenZn335Umg3WhmThKwPE/edit?usp=sharing
